Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-04-05 doc/manuscript_hkk_drought_sensitivity.Rmd
edits to 1st paragraph
@krishnaanujan , I deleted some text with an estimate of C uptake by tropical tree woody growth. This was a very rough, back-of-envelope estimate that I pieced together for the internal grant proposal, but re-reading in this context I realized I'm not comfortable giving that estimate here. (PLus some minor edits.)

<!--KAT suggestion for new intro paragraph (would replace next paragraph)-->
In forests worldwide, intense drought is becoming more frequent, affecting forest demography, diversity, and  C cycling [@intergovernmentalpanelonclimatechangeipcc_climate_2023]. 
Understanding these impacts is an important research priority, particularly for tropical forests, which <!--harbor high species diversity and --> are critical C sinks; tropical tree woody growth removes roughly `r round (150*1/6,0)` Gt CO~2~ annually offsetting ~`r round(7/36.3*100,-1)`% of anthropogenic fossil fuel emissions [@beer_terrestrial_2010;@anderson-teixeira_carbon_2021;@anderson-teixeira_carbon_2016;@brando_droughts_2019].
Drought events can threaten this sink; the 2015-2016 ENSO drought in the Amazon temporarily shifted the region from a net sink to a net source of carbon [@brando_droughts_2019].
However, nuanced study of tropical tree woody growth responses to drought is inhibited by the fact that few tropical species form rings and long-term annual records are rare [@zuidema_tropical_2022]. 
Separate studies have shown that drought characteristics species traits, and microenvironment all affect tree growth responses [@bennett_larger_2015;@mcgregor_tree_2021], but we know little about their combined effects or potential interactions, which could fundamentally alter our understanding of forest responses to drought. 
Specifically, if drought tolerance is influenced independently by drought characteristics, species traits, and microenvironment, tree drought responses may be conserved, and therefore linearly predictable; i.e., a species with low sensitivity to past droughts would have low sensitivity in any microhabitat and through any drought. 
In contrast, if these factors interact, the species and individuals that exhibit resistance to one drought may prove much more vulnerable in a different drought. 
The resultant amplification of the complexity and diversity of drought responses may act as a mechanism for promoting species diversity and stabilizing ecosystem function as drought regimes intensify [@dahinden_future_2017;@luo_tropical_2022;@naumann_global_2018].

>>>>

<!--KAT suggestion for new intro paragraph (would replace next paragraph)-->
In forests worldwide, intense drought is becoming more frequent, affecting forest demography, diversity, and  C cycling [@intergovernmentalpanelonclimatechangeipcc_climate_2023]. 
Understanding these impacts is an important research priority, particularly for tropical forests, whose critical C sink is threatened by drought [@brando_droughts_2019].
However, nuanced study of tropical tree woody growth responses to drought is inhibited by the fact that few tropical species form rings and long-term annual records are rare [@zuidema_tropical_2022]. 
Separate studies have shown that drought characteristics, species traits, and microenvironment all affect tree growth responses [@bennett_larger_2015;@mcgregor_tree_2021], but we know little about their combined effects or potential interactions, which could fundamentally alter our understanding of forest responses to drought. 
Specifically, if drought tolerance is influenced independently by drought characteristics, species traits, and microenvironment, tree drought responses may be conserved, and therefore linearly predictable; i.e., a species with low sensitivity to past droughts would have low sensitivity in any microhabitat and through any drought. 
In contrast, if these factors interact, the species and individuals that exhibit resistance to one drought may prove much more vulnerable in a different drought. 
The resultant amplification of the complexity and diversity of drought responses may act as a mechanism for promoting species diversity and stabilizing ecosystem function as drought regimes intensify [@dahinden_future_2017;@luo_tropical_2022;@naumann_global_2018].
